---
title:  "<span style='font-size: 25px'>RHST vs SDM</style>"
subtitle: "<span style='font-size: 15px'>세종보로 시연, 나머지 하천은 곧바로 래스터 생산에 투입</style>"
author: "전청옥 서울대학교 지리학과"
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    highlight: haddock
    self_contained: yes
    gallery: yes
    number_sections: yes
    pandoc_args: --number-offset=0
    code_folding: show
    toc_depth: 4
    lightbox: yes
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
require(knitr); require(rmdformats); require("DT")

options(width="480"); options(max.print="75")
options(digits=3); options(scipen=1000)
options(DT.options = list(class="display compact nowrap hover",
                          rownames=FALSE));
options(encoding = 'EUC-KR')
knitr::opts_chunk$set(
   echo = TRUE, message = FALSE, warning = FALSE, cache = FALSE,
   comment = ":)  ", collapse = FALSE, prompt = FALSE, tidy = FALSE,
   fig.align="center", fig.retina=2.5,
 aliases=c(h = 'fig.height', w = 'fig.width')
)

knitr::opts_knit$set(width=75)
```   


Effects of eigenfunction-based spatial variables 
on the prediction of Species Distribution Model 

International Biogeography Society conference Vancouver 2022 JAN

# Introduction

To check the effect, here we use dataset from Elith, et al. 2020. which is useful for benchmarking various SDM algorithm.
This dataset(R package disdat) includes 226 anonymised species from 6 regions.
For faster SDM fitting, here use parallel computing(furrr), all species will be seriesd to list.
The process is following.

1. Listing species(P, A, B)
2. Add PCNM variables
3. Modeling
4. Cross-validation
5. Evaluation
6. Interpretation





# 1. Listing species(P, A, B)

```{r}
library(tidyverse)
library(sf)
library(rnaturalearth)
library(disdat)
```

```{r}
getwd()
```

```{r}
my.region <- c("AWT", "CAN", "NSW", "NZ", "SA", "SWI")

SDMdata.AWT <- my.disData("AWT")
SDMdata.CAN <- my.disData("CAN")
SDMdata.NSW <- my.disData("NSW")
SDMdata.NZ <- my.disData("NZ")
SDMdata.SA <- my.disData("SA")
SDMdata.SWI <- my.disData("SWI")

```

```{r}

### AWT #####################################################
SDMdata.AWT.pa <- list()
for(i in unique(SDMdata.AWT$pa$spid)){
  SDMdata.AWT.pa[[i]] <- SDMdata.AWT$pa %>% filter(spid == i)
  
}

SDMdata.AWT.po <- list()
for(i in unique(SDMdata.AWT$po$spid)){
  SDMdata.AWT.po[[i]] <- SDMdata.AWT$po %>% filter(spid == i)
  
}

######## CAN #####################################################
SDMdata.CAN.pa <- list()
for(i in unique(SDMdata.CAN$pa$spid)){
  SDMdata.CAN.pa[[i]] <- SDMdata.CAN$pa %>% filter(spid == i)
  
}

SDMdata.CAN.po <- list()
for(i in unique(SDMdata.CAN$po$spid)){
  SDMdata.CAN.po[[i]] <- SDMdata.CAN$po %>% filter(spid == i)
  
}

################# NSW #######################################
SDMdata.NSW.pa <- list()
for(i in unique(SDMdata.NSW$pa$spid)){
  SDMdata.NSW.pa[[i]] <- SDMdata.NSW$pa %>% filter(spid == i)
  
}

SDMdata.NSW.po <- list()
for(i in unique(SDMdata.NSW$po$spid)){
  SDMdata.NSW.po[[i]] <- SDMdata.NSW$po %>% filter(spid == i)
  
}

############################ NZ #############################
SDMdata.NZ.pa <- list()
for(i in unique(SDMdata.NZ$pa$spid)){
  SDMdata.NZ.pa[[i]] <- SDMdata.NZ$pa %>% filter(spid == i)
  
}

SDMdata.NZ.po <- list()
for(i in unique(SDMdata.NZ$po$spid)){
  SDMdata.NZ.po[[i]] <- SDMdata.NZ$po %>% filter(spid == i)
  
}

####################################### SA ####################
SDMdata.SA.pa <- list()
for(i in unique(SDMdata.SA$pa$spid)){
  SDMdata.SA.pa[[i]] <- SDMdata.SA$pa %>% filter(spid == i)
  
}

SDMdata.SA.po <- list()
for(i in unique(SDMdata.SA$po$spid)){
  SDMdata.SA.po[[i]] <- SDMdata.SA$po %>% filter(spid == i)
  
}

##################################################### SWI #######
SDMdata.SWI.pa <- list()
for(i in unique(SDMdata.SWI$pa$spid)){
  SDMdata.SWI.pa[[i]] <- SDMdata.SWI$pa %>% filter(spid == i)
  
}

SDMdata.SWI.po <- list()
for(i in unique(SDMdata.SWI$po$spid)){
  SDMdata.SWI.po[[i]] <- SDMdata.SWI$po %>% filter(spid == i)
  
}


```

```{r}
SDMdata.AWT.pa$awt01
```
```{r}
map(SDMdata.AWT.pa, function(x){filter(x, pa == 1)})
```

2nd PA doesn't have environmental variables


```{r}
SDMdata.AWT.pa <- map(SDMdata.AWT.pa, function(x){merge(x, SDMdata.AWT$env, by = "siteid")})
SDMdata.CAN.pa <- map(SDMdata.CAN.pa, function(x){merge(x, SDMdata.CAN$env, by = "siteid")}) 
SDMdata.NSW.pa <- map(SDMdata.NSW.pa, function(x){merge(x, SDMdata.NSW$env, by = "siteid")})
SDMdata.NZ.pa <- map(SDMdata.NZ.pa, function(x){merge(x, SDMdata.NZ$env, by = "siteid")})
SDMdata.SA.pa <- map(SDMdata.SA.pa, function(x){merge(x, SDMdata.SA$env, by = "siteid")})
SDMdata.SWI.pa <- map(SDMdata.SWI.pa, function(x){merge(x, SDMdata.SWI$env, by = "siteid")})
```




# 2. Adding PCNM variables

```{r}
library(vegan)
library(geosphere)
```


```{r}
df.ref <- refEllipsoids()

data.frame(region = c("AWT", "CAN", "NSW", "NZ", "SA", "SWI"),
           code = c("RF", "CC", "WE", "IN", "WE", "??")) %>% #ellipsoid code in geosphere::refEllipsoid()
  merge(df.ref, by = "code") %>% 
  mutate(f = 1/invf) -> df.ref2

df.ref2
```

Except NSW and SA, SWI, CAN, NZ, AWT needed to be transfrom to WGS84

load all to sf
```{r}
my.transformer.AWT <-
function(x){
  x %>% 
    sf::st_as_sf(coords = c("x","y"),
                 crs = 28355) %>% 
    sf::st_transform(crs = 4326) %>% 
    st_coordinates() %>%
    cbind(x) %>%
  dplyr::rename(long = X,
                lat = Y)}

my.transformer.CAN <-
function(x){
  x %>% 
    sf::st_as_sf(coords = c("x","y"),
                 crs = 4008) %>% 
    sf::st_transform(crs = 4326) %>% 
    st_coordinates() %>%
    cbind(x) %>%
  dplyr::rename(long = X,
                lat = Y)}

my.transformer.NZ <-
function(x){
  x %>% 
    sf::st_as_sf(coords = c("x","y"),
                 crs = 27200) %>% 
    sf::st_transform(crs = 4326) %>% 
    st_coordinates() %>%
    cbind(x) %>%
  dplyr::rename(long = X,
                lat = Y)}

my.transformer.SWI <-
function(x){
  x %>% 
    sf::st_as_sf(coords = c("x","y"),
                 crs = 21781) %>% 
    sf::st_transform(crs = 4326) %>% 
    st_coordinates() %>%
    cbind(x) %>%
  dplyr::rename(long = X,
                lat = Y)}

sf.AWT.po <- map(SDMdata.AWT.po, my.transformer.AWT)
sf.CAN.po <- map(SDMdata.CAN.po, my.transformer.CAN)
sf.NSW.po <- map(SDMdata.NSW.po, function(x){x %>%
                                              sf::st_as_sf(coords = c("x","y"),
                                                           crs = 4326) %>% 
                                                  st_coordinates() %>%
                                                  cbind(x)} %>% 
                                              dplyr::rename(long = X,
                                                            lat = Y)
                 )
sf.NZ.po <- map(SDMdata.NZ.po, my.transformer.NZ)
sf.SA.po <- map(SDMdata.SA.po, function(x){x %>%
                                              sf::st_as_sf(coords = c("x","y"),
                                                           crs = 4326) %>% 
                                                  st_coordinates() %>%
                                                  cbind(x)} %>% 
                                              dplyr::rename(long = X,
                                                            lat = Y)
                 )

sf.SWI.po <- map(SDMdata.SWI.po, my.transformer.SWI)

```


```{r}
my.geodistm <- function(x){as.dist(geosphere::distm(x[,c("long","lat")], fun = distGeo))}

distm.AWT.po <- map(sf.AWT.po, my.geodistm)
distm.CAN.po <- map(sf.CAN.po, my.geodistm)
distm.NSW.po <- map(sf.NSW.po, my.geodistm)
distm.NZ.po <- map(sf.NZ.po, my.geodistm)
distm.SA.po <- map(sf.SA.po, my.geodistm)
distm.SWI.po <- map(sf.SWI.po, my.geodistm)

```



```{r}
distGeo(SDMdata.AWT.po$awt01, a = df.ref2[4,"a"], f = df.ref2[4,"f"])

```





```{r}
my.checkRegion <- function (region) {
  regions <- c("AWT", "CAN", "NSW", "NZ", "SA", "SWI")
  region <- toupper(region[1])
  if (!region %in% regions) {
    stop("unknown region: ", region, ". Should be one of: ", 
      paste(regions, collapse = ", "))
  }
  region
}

my.reshape_pa <- function (x){
  r <- stats::reshape(x, varying = colnames(x)[-c(1:4)], times = colnames(x)[-c(1:4)], 
    v.names = "pa", timevar = "spid", direction = "long")
  r$id = NULL
  rownames(r) <- NULL
  r
}

my.pkgPath <- 
function (subdir = "") {
  system.file(file.path("extdata", subdir), package = "disdat")
}



my.disData <- function(region){
  region <- my.checkRegion(region)
  path <- my.pkgPath()
  f <- list.files(pattern = region, path, full.names = TRUE)
  nms <- gsub("\\.rds", "", basename(f))
  nms <- gsub(region, "", nms)
  nms <- gsub("test_", "", nms)
  nms <- gsub("train_", "", nms)
  if (length(f) == 4) {
    x <- lapply(f, readRDS)
    names(x) <- nms
    i <- which(nms == "pa")
    x[[i]] <- my.reshape_pa(x[[i]])
  }
  else {
    fe <- grep("_env", f, value = TRUE)
    env <- lapply(fe, readRDS)
    env <- do.call(rbind, env)
    fa <- grep("_pa", f, value = TRUE)
    pa <- lapply(fa, readRDS)
    pa <- lapply(pa, my.reshape_pa)
    pa <- do.call(rbind, pa)
    bg <- readRDS(grep("train_bg", f, value = TRUE))
    po <- readRDS(grep("train_po", f, value = TRUE))
    x <- list(env = env, pa = pa, bg = bg, po = po)
  }
  return(x)
}

my.disData("NSW")



```


```{r}
SDMdata[[1]]$pa
```



```{r}
disPa("AWT")
```













































































```{r}

point.test <- st_as_sf(data.frame(long = rep(c(0, 30, 60, 90, 120, 150, 180, -30, -60, -90, -120, -150), each = 8),
                                  lat = rep(-90 + (c(2,5,9,14,20,27,35,44)/10), 12)),
                       coords = c("long", "lat"),
                       crs = 4326,
                       remove = F) 

point.test <- sf::st_transform(point.test, crs = 3976) 



boundary <- ne_countries(returnclass = "sf") %>% st_transform(3412)

ggplot() +
 geom_sf(data = point.test)
   
```
```{r}
test <- sf::st_transform(point.test, crs = 3976) 
ggplot() +
 geom_sf(data = test)
```


```{r}
round <- sf::st_read("PCNM_in_graphic/round.shp")
df.round <- sf::st_coordinates(round)

```
```{r}
library(geosphere)
geodistm <- function(x){as.dist(geosphere::distm(x[,1:2], fun = distGeo))}
distm.round <- geodistm(df.round)
distm.point <- geodistm(st_drop_geometry(point.test))

library(vegan)

pcnm.round <- vegan::pcnm(distm.round)
pcnm.point <- vegan::pcnm(distm.point)

round %>% cbind(pcnm.round$vectors[,1:20]) -> pcnm.round120 
point.test %>% cbind(pcnm.point$vectors[,1:20]) -> pcnm.point120


ggplot(pcnm.point120, aes(color = PCNM2)) +
 geom_sf() 
```

```{r}
library(ggplot2)
ggplot2::ggplot(pcnm.round120[,4:24]) +
  geom_sf()

```
# Loading data

```{r}
library(disdat)
```


```{r}
disdat::disPa(region = "AWT", group = "plant")
```

